---
title: "Modeling Process Quality"
output: html_notebook
---
# Stem-and-Leaf Plot
  - One of the most useful graphica techniques for summarizing and presenting data.

## Process
  - Assuming each data consists of at least two digits.
  - divide each number into two parts:
      - stem, consisting of one or more of the leading digits
      - leaf, consisting of the remaining digits
  - choose relatively few stems in comparison with the number of observations (between 5 and 20 stems)
  - list the chosen stems along the left-hand margin of the display
  - list beside each stem all leaves corresponding to the observed data values according to the order in which they are encountered in the data set.
```{r}
days <- c(48,41,35,36,37,26,36,46,35,47,35,34,36,42,43,36,56,32,46,30,37,43,17,26,28,27,45,33,22,27,16,22,33,30,24,23,22,30,31,17)
stem(days)
```
# Time Series Plot (Run Chart)
- A plot of data values versus time.
- Called a marginal plot if a histogram/dot plot is attached to the margin of the chart.

```{r}
###################################
# First Attempt, ggExtra approach
###################################
library(ggplot2)
library(ggExtra)
library(grid)
times <- c(1:40)
df <- data.frame(times, days)
p <- ggplot(df,aes(times,days))+ geom_point() +theme_classic()
mp <- ggMarginal(p,type="histogram",margins="y")
grid.newpage()
grid.draw(mp)

################################
# Second Attemp, Layout approach
################################
new_layout <- layout(matrix(c(1,2),nrow=1,ncol=2),widths=c(5,1),heights=c(5,5),TRUE)
par(mar=c(5,4,2,0))
plot(df,xlab="Times",ylab="Days",main="Marginal Plot")
par(mar=c(5,0,2,1))
stripchart(days,method="stack",offset=0.5,vertical=TRUE,axes=FALSE)
```
# Histogram
  - is a more compact summary of data than a stem-and-leaf plot.
  - better suited for larger data set, 75 to 100 or more observations
## Process
  - Divide the range of the data into intervals-- **class intervals, cells, ** or **bins** _(determine upper and lower boundary of each bin)_
      - the number of bins depends on the number of observations and the amount of scatter or dispersion in the data. In general, between 5 to 20 bins is statisfactory in most cases.
      - Number of bins should increase with $n$, number of observations
      - approximately equal to the square root of the number observations often works well.
      - **Sturger's rule**, $h=1+\log_2n$ where $n$ is the sample size.
  - determine axes of the histogram 
      - Use horizontal axis to represent the measurement scale for the data
      - Use the vertical scale to represent the counts, or **frequencies**
      - when the frequencies in each bin are divided by the total number of observations $(n)$, it is called **relative frequencies**
  - draw rectangles over each bin so that the height of each rectangle is proportional to the frequenciy, or relative frequency
```{r}  
thick <- c(438,413,444,468,445,472,474,454,455,449,450,450,450,459,466,470,457,441,450,445,487,430,446,450,456,433,455,459,423,455,451,437,444,453,434,454,448,435,432,441,452,465,466,473,471,464,478,446,459,464,441,444,458,454,437,443,465,435,444,457,444,471,471,458,459,449,462,460,445,437,461,453,452,438,445,435,454,428,454,434,432,431,455,447,454,435,425,449,449,452,471,458,445,463,423,451,440,442,441,439)
hist(thick)

defect <- c(6,1,5,7,8,6,0,2,4,2,5,2,4,4,1,4,1,7,2,3,4,3,3,3,6,3,2,3,4,5,5,2,3,4,4,4,2,3,5,7,5,4,5,5,4,5,3,3,3,12)
hist(defect,12)
```
# Numerical Summary of Data
 - Numerical measures of central tendency and scatter.
 - suppose that $x_1, x_2, ...,x_n$ are observations

## Sample Average
$$\bar{x}=\frac{x_1+x_2+...+x_n}{n}=\frac{\sum_{i=1}^{n}x_i}{n}$$
- arithmetic mean of the $n$ observations
```{r}
mean(thick)
```

## Sample Variance
$$s^2=\frac{\sum_{i=1}^{n}(x_i-\bar{x})^2}{n-1}$$
- measure variability in the sample data
- if there is no variability in the sample, each observation $x_i=\bar{x}$ and the sample variance $S^2=0$

## Sample Standard Deviation
$$s=\sqrt{\frac{\sum_{i=1}^n(x_i-\bar{x})^2}{n-1}}$$
# Box Plot (Box Whisker Plot)
- graphical display for data that simultaneously displays:
      - central tendency
      - spread of variability
      - departure from symmetry
      - identification of outliers
- displays the three quartiles (Q1, Q2, and Q3), the minimum, and the maximum of the data on a rectangle box (vertical or horizontal)

```{r}
diameter <- c(120.5,120.9,120.3,121.3,120.4,120.2,120.1,120.5,120.7,121.1,120.9,120.8)
bp <- boxplot(diameter, horizontal=TRUE, xlab="Hole Diameters (mm)", border=TRUE, frame=FALSE,axes=FALSE)
leg <- as.character(bp$stats)
text(x=bp$stats,y=0.7,labels=leg)
```